---
title: "draft"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---


### Setup: Load libraries 

```{r}
library(tidyverse)
library(here)
library(janitor)
library(jtools)

library(gtsummary)
library(gt)

library(MatchIt) # matching
library(cobalt)  # balance + love plots

library(fixest) # fast fixed effects
library(scales) # plotting

```

# Study Background 

## 1. Data Description
After the code block(s) that read in and clean the data, add some notes covering:

What does each row represent? What is the unit of observation?
What are the key outcome, treatment, and control variables?
Any notable data cleaning decisions (dropped observations, constructed variables, etc.)
Sample size, time period, and geographic scope

```{r}
hsr <- read_csv(here("data/xie2025_simulated_panel.csv")) %>%
  mutate(region = as.factor(region), 
         lat_band = as.factor(lat_band),
         city_scale = as.factor(city_scale), 
         hsr_open_year = na_if(hsr_open_year, Inf)) # change Inf to Na
```
2. Annotated Replication
After each subsequent code block, add some notes covering:

What does the code accomplish? (summary of the procedure or model)
What is the causal identification approach? (e.g., DiD, RDD, IV, matching)
What assumptions must hold for the estimates to be causally valid?
Are there any signs in the output that those assumptions are or are not satisfied?


# Staggered DID - Impact of HSR on Avian Species Richness

```{r}
# Column 1: Economic variables
m1 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 2: + Demographic
m2 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 3: + Infrastructure
m3 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 4: + Environmental
m4 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 5: Full controls (preferred specification)
m5 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct +
    wind_speed + precip_mm + humidity_pct + temperature_c |
    city_id + year,
  data = hsr, cluster = ~city_id
)
```


```{r}
# --- Results table ---
etable(
  m1, m2, m3, m4, m5,
  keep    = "hsr",
  headers = c("Economic", "+ Demographic", "+ Infrastructure", "+ Environmental", "+ Weather (Full)"),
  digits  = 4,
  fitstat = c("n", "r2")
)
```


```{r}
# --- Effect size: coef / SD(outcome), as reported in paper ---
sd_outcome <- sd(hsr$avian_increment_ratio_natl, na.rm = TRUE)
effect_size <- coef(m5)["hsr"] / sd_outcome
cat(sprintf("\nEffect size (col 5): %.1f%% of 1 SD\n", abs(effect_size) * 100))
```

# Event study/ parallel trends plot

```{r}
# event_time is already in the data; bin tails at ±5 to avoid sparse cells
hsr_es <- hsr |>
  mutate(event_time_binned = case_when(
    is.na(event_time) ~ NA_real_,   # never-treated cities (control group)
    event_time < -5   ~ -5,
    event_time >  5   ~  5,
    TRUE              ~ event_time
  ))

# Full controls, two-way FE, clustered SEs
# i(event_time_binned, ref = -3): creates dummies for each relative year,
# omitting -3 as the reference period
es_model <- feols(
  avian_increment_ratio_natl ~
    i(event_time_binned, ref = -3) +
    gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct +
    wind_speed + precip_mm + humidity_pct + temperature_c |
    city_id + year,
  data    = hsr_es,
  cluster = ~city_id
)

# --- Extract coefficients and CIs ---
es_coefs <- broom::tidy(es_model, conf.int = TRUE) |>
  filter(str_detect(term, "event_time_binned")) |>
  mutate(event_time = as.numeric(str_extract(term, "-?\\d+"))) |>
  # Add the reference year back in as zero
  bind_rows(tibble(event_time = -3, estimate = 0, conf.low = 0, conf.high = 0)) |>
  arrange(event_time)

# --- Plot ---
es_plot <- es_coefs |>
  ggplot(aes(x = event_time, y = estimate)) +
  # Shade pre-treatment region
  annotate("rect", xmin = -5.5, xmax = -0.5,
           ymin = -Inf, ymax = Inf, alpha = 0.04, fill = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dotted",
             color = "steelblue", linewidth = 0.7) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, fill = "steelblue") +
  geom_line(color = "steelblue", linewidth = 0.8) +
  geom_point(color = "steelblue", size = 2.5) +
  # Annotate reference year
  annotate("text", x = -3, y = 0.0015, label = "ref", size = 3,
           color = "grey40") +
  # Annotate HSR opening
  annotate("text", x = 0.1, y = max(es_coefs$conf.high) * 0.9,
           label = "HSR opens →", hjust = 0, size = 3.5, color = "steelblue") +
  scale_x_continuous(breaks = -5:5) +
  labs(
    title    = "Event Study: Effect of HSR on Bird Species Increment",
    subtitle = "Coefficients relative to 3 years before HSR opening | 95% CI shaded\nPre-treatment estimates near zero support the parallel trends assumption",
    x        = "Years Relative to HSR Opening",
    y        = "Δ Avian Increment Ratio (National)",
    caption  = "Two-way FE (city + year), full controls, SEs clustered at city level."
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40", size = 10),
    panel.grid.minor = element_blank()
  )

#ggsave("hsr_event_study.png", es_plot, width = 9, height = 5.5, dpi = 200)

# --- Quick numeric check: are pre-treatment estimates jointly zero? ---
# Tests the parallel trends assumption formally
pre_terms <- es_coefs |>
  filter(event_time < -3) |>
  pull(event_time) |>
  (\(x) paste0("event_time_binned::", x))()

if (length(pre_terms) > 0) {
  joint_test <- wald(es_model, pre_terms)
  cat("\nJoint test of pre-treatment coefficients (H0: all zero):\n")
  print(joint_test)
}

```
```{r}
# Register controls shorthand for use in feols formulas
setFixest_fml(..controls = ~ gdp_growth_pct + share_primary_pct + share_secondary_pct +
  population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct +
  wind_speed + precip_mm + humidity_pct + temperature_c)

# Preferred spec (col 1 benchmark) — reference point for all comparisons
m_preferred <- feols(
  avian_increment_ratio_natl ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# =============================================================================
# TABLE 16: Robustness checks on the preferred specification
# =============================================================================

# --- Col 2: Log transformation of national-level DV ---
# Add small constant to handle zeros before logging
m_log <- feols(
  log(avian_increment_ratio_natl + 1e-6) ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# --- Col 3: Replace DV with province-level ratio ---
m_prov <- feols(
  avian_increment_ratio_prov ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# --- Col 4: Log of province-level ratio ---
m_prov_log <- feols(
  log(avian_increment_ratio_prov + 1e-6) ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# --- Col 5: Exclude city-years before HSR ever opens (pre-operation years) ---
# Keep only: never-treated cities (all years) + treated cities from open_year onward
hsr_post_only <- hsr |>
  filter(is.na(hsr_open_year) | year >= hsr_open_year)

m_post_only <- feols(
  avian_increment_ratio_natl ~ hsr + ..controls |
    city_id + year,
  data = hsr_post_only, cluster = ~city_id
)

# --- Col 6: Add province fixed effects ---
m_prov_fe <- feols(
  avian_increment_ratio_natl ~ hsr + ..controls |
    city_id + year + province_id,
  data = hsr, cluster = ~city_id
)

# --- Col 7: Cluster SEs at city level (same as preferred but explicit) ---
m_cluster_city <- feols(
  avian_increment_ratio_natl ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id   # already city-clustered; shown for clarity
)

# --- Col 8: Two-way clustering (city + year) ---
m_cluster_2way <- feols(
  avian_increment_ratio_natl ~ hsr + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id + year
)

# --- Col 9: Control for number of birdwatchers (H) ---
m_birdwatchers <- feols(
  avian_increment_ratio_natl ~ hsr + H + ..controls |
    city_id + year,
  data = hsr, cluster = ~city_id
)

cat("=== Table 16: Robustness Checks ===\n\n")
etable(
  m_preferred, m_log, m_prov, m_prov_log,
  m_post_only, m_prov_fe, m_cluster_city, m_cluster_2way, m_birdwatchers,
  keep    = "hsr",
  headers = c("Preferred", "Log DV", "Prov. ratio", "Log prov. ratio",
              "Excl. pre-op", "Province FEs", "Cluster city",
              "2-way cluster", "Birdwatchers"),
  digits  = 4,
  fitstat = c("n", "r2")
)

# =============================================================================
# TABLE 9: PSM-DID
# Match treated and control cities on pre-treatment covariates,
# then re-run the DID on the matched sample
# =============================================================================

# PSM requires one observation per city (use pre-treatment means as baseline)
# Match on: economic, demographic, infrastructure, environmental, and geography
city_baseline <- hsr |>
  filter(year < min(hsr_open_year, na.rm = TRUE)) |>   # pre-HSR era only
  group_by(city_id) |>
  summarise(
    treated_city       = first(treated_city),
    gdp_growth_pct     = mean(gdp_growth_pct,     na.rm = TRUE),
    share_primary_pct  = mean(share_primary_pct,  na.rm = TRUE),
    share_secondary_pct= mean(share_secondary_pct,na.rm = TRUE),
    population_10k     = mean(population_10k,     na.rm = TRUE),
    urban_road_10k_m2  = mean(urban_road_10k_m2,  na.rm = TRUE),
    pm25               = mean(pm25,               na.rm = TRUE),
    green_coverage_pct = mean(green_coverage_pct, na.rm = TRUE),
    wind_speed         = mean(wind_speed,         na.rm = TRUE),
    precip_mm          = mean(precip_mm,          na.rm = TRUE),
    humidity_pct       = mean(humidity_pct,       na.rm = TRUE),
    temperature_c      = mean(temperature_c,      na.rm = TRUE),
    longitude          = first(longitude),
    latitude           = first(latitude),
    .groups = "drop"
  ) |>
  drop_na()

# Fit propensity score model
psm_formula <- treated_city ~ gdp_growth_pct + share_primary_pct +
  share_secondary_pct + population_10k + urban_road_10k_m2 +
  pm25 + green_coverage_pct + wind_speed + precip_mm +
  humidity_pct + temperature_c + longitude + latitude

# Four matching methods to mirror the paper
run_psm_did <- function(method, ...) {
  match_out <- matchit(
    psm_formula,
    data     = city_baseline,
    method   = method,
    distance = "glm",
    ...
  )

  matched_ids <- match.data(match_out)$city_id
  hsr_matched <- hsr |> filter(city_id %in% matched_ids)

  feols(
    avian_increment_ratio_natl ~ hsr + ..controls |
      city_id + year,
    data = hsr_matched, cluster = ~city_id
  )
}

# k-nearest neighbors (k=3, k=5)
m_nn3    <- run_psm_did("nearest", ratio = 3)
m_nn5    <- run_psm_did("nearest", ratio = 5)

# Radius matching: nearest neighbor within 0.05 SD of propensity score
m_radius <- run_psm_did("nearest", caliper = 0.05)

# Optimal full matching (closest equivalent to kernel matching)
m_kernel <- run_psm_did("full")

cat("\n=== Table 9: PSM-DID ===\n\n")
etable(
  m_preferred, m_nn3, m_nn5, m_radius, m_kernel,
  keep    = "hsr",
  headers = c("Full sample", "k-NN (k=3)", "k-NN (k=5)",
              "Radius", "Kernel (approx)"),
  digits  = 4,
  fitstat = c("n", "r2")
)


```

3. Critical Evaluation
Somewhere in the write-up (or as a closing section), address each of the following:

Causal identification: Is the identification strategy credible? What would have to be true for the estimates to be causal?
Statistical assumptions: Are key assumptions (parallel trends, exclusion restriction, etc.) tested or defended? Do you find the evidence convincing?
Sampling & external validity: Who is in the sample? To whom can results be generalized?
Measurement: Are the outcome and treatment variables measured accurately and without systematic bias?
Other limitations: Anything else that affects how much weight you would put on the study’s conclusions?