---
title: "draft"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---


### Setup: Load libraries 

```{r}
library(tidyverse)
library(here)
library(janitor)
library(jtools)

library(gtsummary)
library(gt)

library(MatchIt) # matching
library(cobalt)  # balance + love plots

library(fixest) # fast fixed effects
library(scales) # plotting

```

# Study Background 

## 1. Data Description
After the code block(s) that read in and clean the data, add some notes covering:

What does each row represent? What is the unit of observation?
What are the key outcome, treatment, and control variables?
Any notable data cleaning decisions (dropped observations, constructed variables, etc.)
Sample size, time period, and geographic scope

```{r}
hsr <- read_csv(here("data/xie2025_simulated_panel.csv")) %>%
  mutate(region = as.factor(region), 
         lat_band = as.factor(lat_band),
         city_scale = as.factor(city_scale), 
         hsr_open_year = na_if(hsr_open_year, Inf)) # change Inf to Na

view(match_hsr)
```
2. Annotated Replication
After each subsequent code block, add some notes covering:

What does the code accomplish? (summary of the procedure or model)
What is the causal identification approach? (e.g., DiD, RDD, IV, matching)
What assumptions must hold for the estimates to be causally valid?
Are there any signs in the output that those assumptions are or are not satisfied?


# Staggered DID - Impact of HSR on Avian Species Richness

```{r}
# Column 1: Economic variables
m1 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 2: + Demographic
m2 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 3: + Infrastructure
m3 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 4: + Environmental
m4 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct |
    city_id + year,
  data = hsr, cluster = ~city_id
)

# Column 5: Full controls (preferred specification)
m5 <- feols(
  avian_increment_ratio_natl ~ hsr + gdp_growth_pct + share_primary_pct + share_secondary_pct +
    population_10k + urban_road_10k_m2 + pm25 + green_coverage_pct +
    wind_speed + precip_mm + humidity_pct + temperature_c |
    city_id + year,
  data = hsr, cluster = ~city_id
)
```


```{r}
# --- Results table ---
etable(
  m1, m2, m3, m4, m5,
  keep    = "hsr",
  headers = c("Economic", "+ Demographic", "+ Infrastructure", "+ Environmental", "+ Weather (Full)"),
  digits  = 4,
  fitstat = c("n", "r2")
)
```


```{r}
# --- Effect size: coef / SD(outcome), as reported in paper ---
sd_outcome <- sd(hsr$avian_increment_ratio_natl, na.rm = TRUE)
effect_size <- coef(m5)["hsr"] / sd_outcome
cat(sprintf("\nEffect size (col 5): %.1f%% of 1 SD\n", abs(effect_size) * 100))
```


3. Critical Evaluation
Somewhere in the write-up (or as a closing section), address each of the following:

Causal identification: Is the identification strategy credible? What would have to be true for the estimates to be causal?
Statistical assumptions: Are key assumptions (parallel trends, exclusion restriction, etc.) tested or defended? Do you find the evidence convincing?
Sampling & external validity: Who is in the sample? To whom can results be generalized?
Measurement: Are the outcome and treatment variables measured accurately and without systematic bias?
Other limitations: Anything else that affects how much weight you would put on the study’s conclusions?